---
title: "DPI 610 Exam Section 3"
subtitle: DPI 610
output:
  pdf_document: default
  html_notebook: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning=FALSE)
library(tidyverse)
library(olsrr)
library(Ecdat)
library(boot)

df <- data.frame(
  year = seq(1952,2016,4),
  gdp_q2 = c(0.9,3.3,-2.1,4.4,6.9,9.4,3,-8,7.1,5.4,4.4,6.8,7.5,3.1,2.1,1.7,1.9),
  inc_poll_pre = c(47.4,49.2,49.2,50.3,34.4,58.2,47.3,79.2,53.2,58.4,40.4,42.9,56.8,46.7,52.9,63.2,51.5),
  inc_poll_post = c(40.9,54.1,47.9,69.8,44.3,64.9,42.3,50,61.1,53,43.8,58.1,54.3,53.1,48.5,51.5,52.6),
  inc_party_voteshare = c(44.6,57.8,49.9,61.3,49.6,61.8,48.9,44.7,59.2,53.9,46.5,54.7,50.3,51.3,46.3,51.9,51.1)
)
```

# A Simple Forecasting Model

This question examines a fundamental model of forecasting U.S. Presidential Elections based on an economic indicator and on polls. The data.frame `df` includes data for forecasting elections, with variables tracking: 

- year of the election (`year`)
- 2nd quarter GDP growth in the year of the election (`gdp_q2`)
- poll results for the incumbent party's presidential nominee (`inc_poll_pre`) _before_ the party conventions
- poll results for the incumbent party's presidential nominee (`inc_poll_post`) _after_ the party conventions
- the presidential election vote share for the incumbent party (`inc_party_voteshare`)

## Question 1: Exploring the Election Data

What is the average (across all years in the data) for the incumbent party in the *post-convention* polling? For Q2 GDP growth? What year had the most lopsided election in terms of two-party voteshare since 1952?

## Answer 1

```{r q1}
mean(df$inc_poll_post)
mean(df$gdp_q2)
df$year[which(grepl(max(df$inc_party_voteshare), df$inc_party_voteshare))]
```

The average (across all years in the data) for the incumbent party in the *post-convention* polling is 52.36471%. 
The average (across all years in the data) for Q2 GDP growth is 3.4. 
The year 1972 had the most lopsided election in terms of two-party voteshare since 1952.

## Question 2: Estimating an Election Model

Using an ordinary least squares regression for the years 1952 to 2016, estimate the relationship between incumbent party vote share and the predictors Q2 GDP and pre-conventions polling. Interpret the relationship between vote share and each of these variables. 

## Answer 2

```{r q2}
olsr_pre <- ols_regress(inc_party_voteshare ~ inc_poll_pre + gdp_q2, data = df)
olsr_pre
```

For an ordinary least squares regression for the years 1952 to 2016, the relationship between incumbent party vote share and the predictors Q2 GDP and pre-conventions polling is characterized by the following:
* Holding the Q2 GDP variable constant, for every four years (or every election cycle), pre-conventions polling affects incumbent party vote share by 0.198%, with a standard error of 0.126%.
* Holding the pre-conventions polling variable constant, for every four years (or every election cycle), Q2 GDP affects incumbent party vote share by 1.020%, with a standard error of 0.306%.
This model is only of moderate positive strength, as we're observing R-Squared values of less than 0.50; in addition, for the pre-conventions polling variable, the lower and upper bounds for the standardized test statistic score enclose the 0 value, meaning the variable's effect is not statistically significant.

## Question 3: Re-Estimating an Election Model

Again, using an ordinary least squares regression for the years 1952 to 2016, estimate the relationship between incumbent party vote share and the predictors Q2 GDP and *post-convention* polling. Note that the only difference between the model in this question and the previous one is replacing pre-convention polling with post-convention polling as an explanatory variable. Interpret the relationship between vote share and each of these variables. What accounts for the different results in the two models? 

## Answer 3

```{r q3}
olsr_post <- ols_regress(inc_party_voteshare ~ inc_poll_post + gdp_q2, data = df)
olsr_post
```

For an ordinary least squares regression for the years 1952 to 2016, the relationship between incumbent party vote share and the predictors Q2 GDP and post-conventions polling is characterized by the following:
* Holding the Q2 GDP variable constant, for every four years (or every election cycle), post-conventions polling affects incumbent party vote share by 0.530%, with a standard error of 0.081%.
* Holding the pre-conventions polling variable constant, for every four years (or every election cycle), Q2 GDP affects incumbent party vote share by 0.372%, with a standard error of 0.155%.
This model is of strong positive strength, as we're observing R-Squared values greater than 0.75; in addition, the standard error for post-convention polling is comparatively smaller than that of the pre-convention polling variable, its lower and upper bounds do not enclose 0, and its coefficient is more than twice that of the pre-convention polling variable. Because of these differences in particular, we can infer that post-convention polling is a more accurate indicator of incumbent party vote share.

## Question 4: Out-of-Sample Forecasts

Suppose you would like to assess the performance for each of the two possible models from the previous two questions by performing *out-of-sample predictions* for each election year (e.g., for each year election year, (1) fit a model using _all other election years_, and (2) then make a prediction for the election year in question). This is known as leave-one-out cross validation. What is the mean absolute error for the predictions made by each of the two models? In terms of identifying the national popular-vote winner, for what share of elections does each model make the right prediction? 

```{r q4}
loocv <- function(m, y) {
  x1 <- df %>% 
    filter(year == y)
  
  x2 <- df %>% 
    filter(year != y)
  
  if(m == "pre") {
    folsr <- ols_regress(inc_party_voteshare ~ inc_poll_pre + gdp_q2, data = x2)
  }
  
  if(m == "post"){
    folsr <- ols_regress(inc_party_voteshare ~ inc_poll_post + gdp_q2, data = x2)
  }
  
  as.numeric(predict(folsr$model, newdata = x1)[1])
}

for(i in 1:nrow(df)) {
  df$pre[i] <- round(loocv("pre", df$year[i]), digits = 1)
  df$post[i] <- round(loocv("post", df$year[i]), digits = 1)
  df$prediff[i] <- df$pre[i] - df$inc_party_voteshare
  df$postdiff[i] <- df$post[i] - df$inc_party_voteshare
}

df <- df %>% 
  mutate(actual = ifelse(inc_party_voteshare > 50, 1, 0),
         preres = ifelse(pre > 50, 1, 0),
         postres = ifelse(post > 50, 1, 0),
         preacc = ifelse(actual == preres, 1, 0),
         postacc = ifelse(actual == postres, 1, 0))

df %>% 
  summarize(premae = mean(prediff),
            postmae = mean(postdiff),
            preacc = mean(preacc),
            postacc = mean(postacc))
```

The mean absolute error for the predictions made by the model with pre-convention polling was 7.541176.
The mean absolute error for the predictions made by the model with post-convention polling was 7.505882.
In terms of identifying the national popular-vote winner, the model with pre-convention polling makes the right prediction 76.47059% of the time, according to this dataset.
In terms of identifying the national popular-vote winner, the model with post-convention polling makes the right prediction 100% of the time, according to this dataset.